### 2020-05-17 writeup/proof_transfer_ex.tex
sec 3.3

\subsection{Proofs of Proposition \ref{prop_covariate}}

@@

In Figure \ref{fig_te_complement}, we plot the test error of the target task for $n_2 = 4p$ and $n_1$ ranging from $p$ to $20p$. First we notice that the curves in Figure \ref{fig_te_complement} all cross at the point $n_1=n_2$. In fact, if $n_1=n_2$, then it is easy to observe that $a_1=a_2=(1-\gamma)/2$ is the solution to equation \eqref{compleeq}, where we denote $ \gamma=p/(n_1+n_2)$. Then for any $\lambda$, the test error in \eqref{testcomple} takes the value
$$te(\lambda)= \frac{\gamma}{2}\frac{1}{(1-\gamma)/2}=\frac{p}{n_1+n_2-p}.$$

Second, we observe the following two phases as we increase $n_1 / p$.
\begin{itemize}
	\item When $n_1 \le n_2$, having complementary covariance matrices leads to lower test error compared to the case when $\Sigma_1 = \Sigma_2$.
	\item When $n_1 > n_2$, having complementary covariance matrices leads to higher test error compared to the case when $\Sigma_1 = \Sigma_2$.
\end{itemize}
%Second, from Figure \ref{fig_te_complement} we observe that the complementary cases with $\lambda>1$ is better than the case without covariate shift (i.e. $M=\id$ case) when $n_1<n_2$. On the other hand, if we have enough source task data such that $n_1>n_2$, then it is always better to have no covariate shift.
This phenomenon can be also explained using our theory. With \eqref{compleeq}, we can write
$$te(\lambda)=\frac{\gamma}{2} \cdot \bigbrace{\frac{1}{\frac{a_1}{\lambda^2} + (1-\gamma - a_1)} + \frac{1}{a_1\lambda^2 + (1-\gamma - a_1)}}.$$
We can compute that
\begin{align*}
te(\lambda) - te(1)&= \frac{\gamma}{2(1-\gamma)} (\lambda^2-1)a_1\cdot \bigbrace{  \frac{1}{ -a_1(\lambda^2-1)+(1-\gamma)\lambda^2 } - \frac{1}{a_1(\lambda^2-1) + (1-\gamma)}} \\
&= \frac{\gamma}{2(1-\gamma)} (\lambda^2-1)^2 a_1\cdot  \frac{2a_1 - (1-\gamma) }{[-a_1(\lambda^2-1)+(1-\gamma)\lambda^2 ][a_1(\lambda^2-1) + (1-\gamma)]} .
\end{align*}
If $n_1>n_2$, we have $a_1>(1-\gamma)/2$ (because $a_1>a_2$ as observed from the equation \eqref{compleeq}), and hence $te(\lambda)>te(1)$. Otherwise if $n_1< n_2$, we have $a_1< (1-\gamma)/2$, and hence $te(\lambda)< te(1)$.

>>>>

\subsection{Proofs of Proposition \ref{prop_covariate}}\label{app_covariate}

@@

First we notice that the curves in Figure \ref{fig_te_complement} all cross at the point $n_1=n_2$. In fact, if $n_1=n_2$, then it is easy to observe that $a_1=a_2=(1-\gamma)/2$ is the solution to equation \eqref{compleeq}, where we denote $ \gamma=p/(n_1+n_2)$. Then for any $\lambda$, the test error in \eqref{testcomple} takes the value
$$te(\lambda)= \frac{\gamma}{2}\frac{1}{(1-\gamma)/2}=\frac{p}{n_1+n_2-p}.$$

%Second, from Figure \ref{fig_te_complement} we observe that the complementary cases with $\lambda>1$ is better than the case without covariate shift (i.e. $M=\id$ case) when $n_1<n_2$. On the other hand, if we have enough source task data such that $n_1>n_2$, then it is always better to have no covariate shift.
This phenomenon can be also explained using our theory. With \eqref{compleeq}, we can write
$$te(\lambda)=\frac{\gamma}{2} \cdot \bigbrace{\frac{1}{\frac{a_1}{\lambda^2} + (1-\gamma - a_1)} + \frac{1}{a_1\lambda^2 + (1-\gamma - a_1)}}.$$
We can compute that
\begin{align*}
te(\lambda) - te(1)&= \frac{\gamma}{2(1-\gamma)} (\lambda^2-1)a_1\cdot \bigbrace{  \frac{1}{ -a_1(\lambda^2-1)+(1-\gamma)\lambda^2 } - \frac{1}{a_1(\lambda^2-1) + (1-\gamma)}} \\
&= \frac{\gamma}{2(1-\gamma)} (\lambda^2-1)^2 a_1\cdot  \frac{2a_1 - (1-\gamma) }{[-a_1(\lambda^2-1)+(1-\gamma)\lambda^2 ][a_1(\lambda^2-1) + (1-\gamma)]} .
\end{align*}
If $n_1>n_2$, we have $a_1>(1-\gamma)/2$ (because $a_1>a_2$ as observed from the equation \eqref{compleeq}), and hence $te(\lambda)>te(1)$. Otherwise if $n_1< n_2$, we have $a_1< (1-\gamma)/2$, and hence $te(\lambda)< te(1)$.

### 2020-05-17 writeup/transfer.tex
sec 3.3

\begin{example}
	To compare different choices of $M$ on the performance of $\hat{\beta}_t^{\MTL}$, we assume an upper bound on the scale of $M$.
	Consider the following family of matrices
	\begin{align*}
		\cS_{\mu}\define\bigset{M \mid \det\bigbrace{\frac{1}{p}M^\top M} \le \mu, \lambda(M) \in [\mu_{\min}, \mu_{\max}]},
	\end{align*}
	where $\mu, \mu_{\min}, \mu_{\max}$ are fixed values that do not grow with $p$.
	For the task models, we assume that $\beta_2$ has i.i.d. entries with mean zero and variance $\kappa^2$ and $\beta_1 - \beta_2$ has i.i.d. entries with mean $0$ and variance $d^2$.
	The following proposition shows that when $n_1$ is large enough compared to $n_2$, $\te(\hat{\beta}^{\MTL})$ is minimized within the family of $\cS_{\mu}$ when $M = \mu\id$.
\end{example}

@@

\begin{example}\label{ex_covariate}
	We now consider another case when $\Sigma_1$ and $\Sigma_2$ have complementary eigenspaces. Suppose $\Sigma_1$ and $\Sigma_2$ have the eigendecomposition
$$\Sigma_1^{1/2} = 1+ U\Lambda U^\top, \quad \Sigma_2^{1/2} = 1+ V\Lambda V,$$
where
$$\Lambda = \diag(\wt\lambda_1,\cdots, \wt\lambda_{p/2}), \quad U= (u_1,\cdots, u_{p/2}), \quad V= (v_1,\cdots, v_{p/2}).$$
If $V=U_\perp$, i.e. the vectors $v_1,\cdots, v_{p/2}$ are perpendicular to the vectors $u_1,\cdots, u_{p/2}$, then
$$M=\Sigma_1^{1/2} \Sigma_2^{-1/2}=(1+\Lambda)UU^\top + (1+\Lambda)^{-1}V V^\top .$$

As a concrete example we consider the case where $\wt \lambda_1=\cdots=\wt\lambda_{p/2}$ and we denote $\lambda:=1+\wt\lambda_1$. Thus for $M$, the first $p/2$ singular values are equal to $\lambda$ and the rest ones are equal to $\lambda^{-1}$.
\end{example}

>>>>

\begin{example}\label{ex_cov_family}
	To compare different choices of $M$ on the performance of $\hat{\beta}_t^{\MTL}$, we assume an upper bound on the scale of $M$.
	Consider the following family of matrices
	\begin{align*}
		\cS_{\mu}\define\bigset{M \mid \det\bigbrace{\frac{1}{p}M^\top M} \le \mu, \lambda(M) \in [\mu_{\min}, \mu_{\max}]},
	\end{align*}
	where $\mu, \mu_{\min}, \mu_{\max}$ are fixed values that do not grow with $p$.
	For the task models, we assume that $\beta_2$ has i.i.d. entries with mean zero and variance $\kappa^2$ and $\beta_1 - \beta_2$ has i.i.d. entries with mean $0$ and variance $d^2$.
	The following proposition shows that when $n_1$ is large enough compared to $n_2$, $\te(\hat{\beta}^{\MTL})$ is minimized within the family of $\cS_{\mu}$ when $M = \mu\id$.
\end{example}

@@

\begin{example}\label{ex_covariate}
	Within the setting of Example \ref{ex_cov_family}, we compare two cases: (i) when $M = \id$; (ii) when $M$ has $p/2$ singular values that are equal to $\lambda$ and $p/2$ singular values that are equal to $1 / \lambda$.
	For simplicity, we assume that $d = 0$.
	Hence the two tasks have the same model parameters.
%	now consider another case when $\Sigma_1$ and $\Sigma_2$ have complementary eigenspaces. Suppose $\Sigma_1$ and $\Sigma_2$ have the eigendecomposition
%$$\Sigma_1^{1/2} = 1+ U\Lambda U^\top, \quad \Sigma_2^{1/2} = 1+ V\Lambda V,$$
%where
%$$\Lambda = \diag(\wt\lambda_1,\cdots, \wt\lambda_{p/2}), \quad U= (u_1,\cdots, u_{p/2}), \quad V= (v_1,\cdots, v_{p/2}).$$
%If $V=U_\perp$, i.e. the vectors $v_1,\cdots, v_{p/2}$ are perpendicular to the vectors $u_1,\cdots, u_{p/2}$, then
%$$M=\Sigma_1^{1/2} \Sigma_2^{-1/2}=(1+\Lambda)UU^\top + (1+\Lambda)^{-1}V V^\top .$$
%	As a concrete example we consider the case where $\wt \lambda_1=\cdots=\wt\lambda_{p/2}$ and we denote $\lambda:=1+\wt\lambda_1$. Thus for $M$, the first $p/2$ singular values are equal to $\lambda$ and the rest ones are equal to $\lambda^{-1}$.

	In Figure \ref{fig_te_complement}, we plot the test error of the target task for $n_2 = 4p$ and $n_1$ ranging from $p$ to $20p$.
	Second, we observe the following two phases as we increase $n_1 / p$.
	\begin{itemize}
		\item When $n_1 \le n_2$, having complementary covariance matrices leads to lower test error compared to the case when $\Sigma_1 = \Sigma_2$.
		\item When $n_1 > n_2$, having complementary covariance matrices leads to higher test error compared to the case when $\Sigma_1 = \Sigma_2$.
	\end{itemize}
\end{example}

A theoretical justification of Example \ref{ex_covariate} can be found in Appendix \ref{app_covariate}.
